% Author Alfredo Sánchez Alberca (asalber@ceu.es)

\section{Descriptive Statistics}
\begin{enumerate}[leftmargin=*]
\item Classify the following variables
\begin{enumerate}
\item Daily hours of exercise.
\item Nationality.
\item Blood pressure.
\item Severity of illness. 
\item Number of sport injuries in a year.
\item Daily calorie intake.
\item Size of clothing.
\item Subjects passed in a course. 
\end{enumerate}

\item \label{soccer-injuries}The number of injuries suffered by the members of a soccer team in a league were
\begin{center}
0 \quad 1 \quad 2 \quad 1 \quad 3 \quad 0 \quad 1 \quad 0 \quad 1 \quad 2 \quad 0 \quad 1 \\
1 \quad 1 \quad 2 \quad 0 \quad 1 \quad 3 \quad 2 \quad 1 \quad 2 \quad 1 \quad 0 \quad 1
\end{center}

\begin{enumerate}
\item Construct the frequency distribution table of the sample.
\item Draw the bar chart of the sample and the polygon.
\item Draw the cumulative frequency bar chart and the polygon. 
\end{enumerate}

\item A survey about the daily number of medicines consumed by people over 70 years, shows the following results:

\begin{center}
3\quad 1\quad 2\quad 2\quad 0\quad 1\quad 4\quad 2\quad 3\quad 5\quad 1\quad 3\quad 2\quad 3\quad 1\quad 4\quad 2\quad 4\quad 3\quad 2 \\
3\quad 5\quad 0\quad 1\quad 2\quad 0\quad 2\quad 3\quad 0\quad 1\quad 1\quad 5\quad 3\quad 4\quad 2\quad 3\quad 0\quad 1\quad 2\quad 3
\end{center} 

\begin{enumerate}
\item Construct the frequency distribution table of the sample.
\item Draw the bar chart of the sample and the polygon.
\item Draw the cumulative relative frequency bar chart and the polygon. 
\end{enumerate}

\item In survey about the dependency of older people, 23 persons over 75 years were asked about the help they need in
daily life. 
The answers were
\begin{center}
B\quad D\quad A\quad B\quad C\quad C\quad B\quad C\quad D\quad E\quad A\quad B\quad C\quad E\quad A\quad B\quad C\quad
D\quad B\quad B\quad A\quad A\quad B 
\end{center}

where the meanings of letters are: 
\begin{enumerate}
\item[A] No help.
\item[B] Help climbing stairs.
\item[C] Help climbing stairs and getting up from a chair or bed.
\item[D] Help climbing stairs, getting up and dressing.
\item[E] Help for almost everything.  
\end{enumerate}

Construct the frequency distribution table and the suitable chart.

\item\label{emergency-service} The number of people treated in the emergency service of a hospital every day of November
was
\begin{center}
15 \quad 23 \quad 12 \quad 10 \quad 28 \quad 7 \quad 12 \quad 17 \quad 20 \quad 21 \quad 18 \quad 13 \quad 11 \quad 12 \quad 26 \\
30 \quad 6 \quad 16 \quad 19 \quad 22 \quad 14 \quad 17 \quad 21 \quad 28 \quad 9 \quad 16 \quad 13 \quad 11 \quad 16 \quad 20
\end{center}

\begin{enumerate}
\item Construct the frequency distribution table of the sample.
\item Draw a suitable chart for the frequency distribution. 
\item Draw a suitable chart for the cumulative frequency distribution.
\end{enumerate}

\item The following frequency distribution table represents the distribution of time (in min) required by people
attended in a medical dispensary. 
\[
\begin{array}{|c|c|c|c|c|}
\hline \mbox{Time} & n_{i} & f_{i} & N_{i} & F_{i}\\
\hline 
\left[ 0,5\right) & 2 &  &  &  \\ 
\hline 
\left[ 5,10\right) &  &  & 8 &  \\ 
\hline 
\left[ 10,15\right) &  & &  & 0.7 \\ 
\hline 
\left[ 15,20\right) & 6 &  &  &\\ 
\hline
\end{array}
\]

\begin{enumerate}
\item Complete the table.
\item Draw the ogive. 
\end{enumerate}

\item Use the data of exercise~\ref{soccer-injuries} to calculate the following statistics and interpret them.
\begin{enumerate}
\item Mean.
\item Median.
\item Mode.
\item Quartiles.
\item Percentile 32.
\end{enumerate}

\item The chart below shows the cumulative distribution of the time (in min) required by 20 students to do an exam.
\begin{center}
\resizebox{0.7\textwidth}{!}{\input{img/descriptive/time_exam_ogive}}
\end{center}

\begin{enumerate}
\item A which time have finished half of the students? And 90\% of students?
\item Which percentage of students have finished after 100 minutes?
\item Which is the time that best represent the time required by students in the sample to finish the exam? Is this
value representative or not?
\end{enumerate}

\item In a study about the children growth two samples where drawn, one for newborns and the other for one year old. 
The height in cm of children in both samples were 
\begin{center}
\begin{tabular}{rl}
Newborn children: & 51, 50, 51, 53, 49, 50, 53, 50, 47, 50\\
One year old children: & 62, 65, 69, 71, 65, 66, 68, 69
\end{tabular}
\end{center}

In which group is more representative the mean? Justify the answer.

\item To determine the accuracy of a method for measuring hematocrit in blood, the measurement was repeated 8 times on
the same blood sample.
The results in percentage of hematocrit in plasma were
\[
42.2\quad 42.1\quad 41.9\quad 41.8\quad 42\quad 42.1\quad 41.9\quad 42
\]
What do you think about the accuracy of the method?


\item The histogram below shows the frequency distribution of the body mass index (BMI) of a group of people by gender.
\begin{center}
\resizebox{0.6\textwidth}{!}{\input{img/descriptive/bmi_gender_histogram}}
\end{center} 

\begin{enumerate}
\item Draw the pie chart for the gender.
\item In which group is more representative the mean of the BMI?
\item Calculate the mean for the whole sample.
\end{enumerate}
Use the following sums\\
Males: $\sum x_i=1002$ kg/m$^2$ \quad $\sum x_i^2 = 22781$ kg$^2$/m$^4$ \\
Females: $\sum x_i=1160$ kg/m$^2$ \quad $\sum x_i^2 = 29050$ kg$^2$/m$^4$ 

\item The following table represents the frequency distribution of the yearly uses of a health insurance in a sample of
clients of a insurance company.

\begin{center}
\begin{tabular}{lrrrrrrr}
\toprule
Uses: & 0 & 1 & 2 & 3 & 4 & 5 & 7 \\
Clients: & 4 & 8 & 6 & 3 & 2 & 1 & 1\\
\bottomrule
\end{tabular}
\end{center}

Draw the box plot. How is the symmetry of the distribution?


\item The box plots below correspond to the age of a sample of people by marital status.
\begin{center}
\resizebox{0.6\textwidth}{!}{\input{img/descriptive/age_marital_status_boxplot}}
\end{center} 

\begin{enumerate}
\item Which group has higher ages?
\item Which group has lower central dispersion?
\item Which groups have outliers?
\item Which group has a distribution of ages more asymmetric?
\end{enumerate}


\item The following table represents the frequency distribution of ages at which a group of people suffered a heart
attack. 
\begin{center}
\begin{tabular}{lccccc}
\toprule
Age & [40-50) & [50-60) & [60-70) & [70-80) & [80-90)  \\ 
Persons & 6 & 12 & 23 & 19 & 5  \\ 
\bottomrule
\end{tabular}
\end{center}

Could we assume that the sample comes from a normal population?

Use the following sums: $\sum x_i= 4275$ years, $\sum (x_i-\bar x)^2=7462$ years$^2$, $\sum (x_i-\bar x)^3=-18249$
years$^3$, $\sum (x_i-\bar x)^4=2099636$ years$^4$.

El 
\begin{tabular}{|l|l|l|}
\hline
\multicolumn{1}{|c|}{Nº Sesiones} & \multicolumn{1}{c|}{Tratamiento A} & \multicolumn{1}{c|}{Tratamiento B} \\
\hline
\multicolumn{1}{|c|}{20-40} & \multicolumn{1}{c|}{5} & \multicolumn{1}{c|}{8} \\
\hline
\multicolumn{1}{|c|}{40-60} & \multicolumn{1}{c|}{20} & \multicolumn{1}{c|}{15} \\
\hline
\multicolumn{1}{|c|}{60-80} & \multicolumn{1}{c|}{18} & \multicolumn{1}{c|}{20} \\
\hline
\multicolumn{1}{|c|}{80-100} & \multicolumn{1}{c|}{7} & \multicolumn{1}{c|}{7} \\
\hline
\end{tabular}





\end{enumerate}