% Author Alfredo Sánchez Alberca (asalber@ceu.es)

\section{Discrete random variables}
\begin{enumerate}[leftmargin=*,resume]

\item Let $X$ be a discrete random variable with the following probability distribution
\[
\begin{array}{|c|c|c|c|c|c|}
\hline
X & 4 & 5 & 6 & 7 & 8 \\ 
\hline
f(x) & 0.15 & 0.35 & 0.10 & 0.25 & 0.15 \\ 
\hline
\end{array}
\]
\begin{enumerate}
\item  Calculate and represent graphically the distribution function.
\item  Calculate the following probabilities
\begin{enumerate}
\item  $P(X<7.5)$.
\item  $P(X>8)$.
\item  $P(4\leq X\leq 6.5)$.
\item  $P(5<X<6)$.
\end{enumerate}
\end{enumerate}

\item Let $X$ be a discrete random variable with the following probability distribution
\[
F(x)=
\begin{cases}
0 & \text{si $x<1$,} \\
1/5 & \text{si $1\leq x< 4$,} \\
3/4 & \text{si $4\leq x<6$,} \\
1 & \text{si $6\leq x$.}
\end{cases}
\]

\begin{enumerate}
\item  Calculate the probability function.
\item  Calculate the following probabilities
\begin{enumerate}
\item  $P(X=6)$.
\item  $P(X=5)$.
\item  $P(2<X<5.5)$.
\item  $P(0\leq X<4)$.
\end{enumerate}
\item Calculate the mean. 
\item Calculate the standard deviation. 
\end{enumerate}

\item An experiment consist in injecting a virus to three different types of rats and observing if they survive or not. 
It is known that the probability of surviving is $0.5$ for the first type of rats, $0.4$ for the second type and $0.3$
for the third type.
\begin{enumerate}
\item Calculate the probability function of the variable $X$ that measures the number of surviving rats.
\item Calculate the distribution function.
\item Calculate $P(X\leq 1)$, $P(X\geq 2)$ y $P(X=1.5)$.
\item Calculate the mean and the standard deviation. 
Is representative the mean?
\end{enumerate}

\iem The chance of being cured with a treatment is $0.85$. 
If we apply the treatment to 6 patients,
\begin{enumerate}
\item What is the probability that half of them are cured?
\item What is the probability that a least 4 of them are cured?
\end{enumerate}

\item Ten persons came into contact with a person infected by tuberculosis. 
The probability of being infected after contacting a person with tuberculosis is $0.10$.
\begin{enumerate}
\item What is the probability that nobody are infected?
\item What is the probability that at least 2 persons are infected?
\item What is the expected number of infected persons? 
\end{enumerate}

\item The probability of having an adverse reaction to a vaccine is $0.001$. 
If 2000 persons are vaccinated, what is the probability of having some adverse reaction?

\item The average number of calls per minute that arrive to a telephone switchboard is 120. 
\begin{enumerate}
\item What is the probability of receiving less than 4 calls in 2 seconds?
\item What is the probability of receiving at least 3 calls in 3 seconds?
\end{enumerate}

\item A test contains 10 questions with 3 possible options each. 
For every question you win a point if you give the right answer and loss half a point if the answer is wrong. 
A student knows the right answer for 3 of the 10 questions and answers the rest randomly. 
What is the probability of passing the exam?

\item It has been observed experimentally that 1 of every 20 trillions of cells exposed to radiation mutates becoming
carcinogenic. 
We know that the human body has approximately 1 trillion of cells by kilogram ot tissue. 
Calculate the probability that a 60 kg person exposed to radiation develops cancer.
If the radiation affects 3 persons weighing 60 kg, what is the probability that a least one of them develops cancer?

\item A diagnostic test for a disease returns 1\% of positive outcomes, and the positivie and negative predictive values
are $0.95$ and $0.98$ respectively. 
\begin{enumerate}
\item Calculate the prevalence of the disease.
\item Calculate the sensitivity and the specificity of the test.
\item If the test is applied to 12 sick persons, what is the probability of getting at least a wrong diagnosis?
\item If the test is applied to 12 persons, what is the probability of getting a right diagnosis for all of them?
\end{enumerate}

\item In a study about a parasite that attack the kidney of rats it is known that the average number of parasites per
kidney is 3. 
\begin{enumerate}
\item Calculate the probability that a rat have more than 3 parasites.
\item Calculate the probability of having at least 9 rats infected in a sample of 10 rats. 
\end{enumerate}

\item In a physiotherapy course there are 60\% of females and 40\% of males.
\begin{enumerate}
\item If 6 random students have to go to a hospital for making practices, what is the probability of going more males
than females?
\item In 5 samples of 6 students, what is the probability of having some sample without males?
\end{enumerate}

\end{enumerate}


