% Author Alfredo Sánchez Alberca (asalber@ceu.es)

\section{Probability}
\begin{enumerate}[leftmargin=*,resume]
\item Construct the sample space of the following random experiments:
\begin{enumerate}
\item Pick a random person and measure the gender and whether she or he is smoker or not. 
\item Pick a random person and measure the blood type and whether she or he is smoker or not.
\item Pick a random person and measure the gender, the blood type and whether she or he is smoker or not.
\end{enumerate}

\item There are two boxes with balls of different colors. 
The first box contains 3 white balls and 2 black balls, and the second one contains 2 blue balls, 1 red ball and 1 green
ball. 
Construct the sample space of the following random experiments:
\begin{enumerate}
\item Pick a random ball from every box. 
\item Pick two random balls from every box.
\end{enumerate}

\item The Morgan's laws state that given two events $A$ and $B$ from the same sample space, $\overline{A\cup B}=\bar A
\cap \bar B$ and $\overline{A\cap B}=\bar A \cup \bar B$.
Proof both statements graphically using Venn diagrams.

\item Calculate the probability of the following events of the random experiment consisting in tossing 3 coins: 
\begin{enumerate}
\item Get exactly 1 heads. 
\item Get exactly 2 tails.
\item Get two or more heads.
\item Get some tails. 
\end{enumerate}

\item In a laboratory there are 4 flasks with sulfuric acid and 2 with nitric acid, and in another laboratory there are
1 flask with sulfuric acid and 3 with nitric acid. 
A random experiment consist in picking two flask, one from every laboratory. 
Calculate the probability of the following events:

\begin{enumerate}
\item The two picked flasks are of sulfuric acid.
\item The two picked flasks are of nitric acid.
\item The tow picked flasks contains different acids.
\end{enumerate}
Calculate the same probabilities if the flask picked in the first laboratory is put in the second laboratory before
picking the flask from it. 

\item Let $A$ and $B$ be events of the same sample space, such that $P(A)=3/8$, $P(B)=1/2$, $P(A\cap B)=1/4$.
Calculate the following probabilities:
\begin{enumerate}
\item  $P(A\cup B)$.
\item  $P(\bar A)$ y $P(\bar B)$.
\item  $P(\bar A\cap \bar B)$.
\item  $P(A\cap \bar B)$.
\item  $P(A|B)$.
\item  $P(A|\bar B)$.
\end{enumerate}

\item In a hospital the probability of getting hepatitis in a blood transfusion from a unit of blood is $0.01$.
A patient gets two units of blood while staying at the hospital.
What is the probability of getting hepatitis?

\item Let $A$ and $B$ be two events of the same sample space, such that $P(A)=0.6$ and $P(A\cup B)=0.9.$
Calculate $P(B)$ with the following assumptions:
\begin{enumerate}
\item $A$ and $B$ are incompatible.
\item $A$ and $B$ are independent.
\end{enumerate}

\item A study about smoking has published that 40\% of smokers have a smoker father, 25\% have a smoker mother and 52\%
have al least one of the parents smoker.
We pick a random person from this population.
Answer the following questions: 
\begin{enumerate}
\item What is the probability of having a smoker mother if the father smokes?
\item What is the probability of having a smoker mother if the father doesn't smoke?
\item Are independent the events having a smoker father and having a smoker mother?
\end{enumerate}

\item The probability that an injury $A$ is repeated is $4/5$, the probability that another injury $B$ is repeated is
$1/2$, and the probability that both injuries are repeated is $1/3$.
Calculate the probability of the following events:
\begin{enumerate}
\item Only injury $B$ is repeated.
\item At least one injury is repeated.
\item Injury $B$ is repeated if injury $A$ has been repeated.
\item Injury $B$ is repeated if injury $A$ hasn't been repeated. 
\end{enumerate}

\item In a digestive clinic from every 1000 patients that arrive with stomach pain, 700 have gastritis,
200 have an ulcer and 100 have cancer.
After analyzing the gastric symptoms, it is known that the probability of having vomiting is $0.3$ in case of gastritis,
$0.6$ in case of ulcer and $0.9$ in case of cancer. 
What is the diagnosis for a new patient with stomach pain that has vomiting?

Note: Assume that the only diseases are gastritis, ulcer and cancer and that are incompatible among them.  

\item To evaluate the effectiveness of a diagnosis test, the test was applied to a sample of 6900 people with the
following results:

\begin{center}
\begin{tabular}{|l|c|c|}
\cline{2-3}
\multicolumn{1}{l|}{} & Test $+$ & Test $-$ \\
\hline
Sick & 4680 & 120 \\
\hline
Healthy & 120 & 2020 \\
\hline
\end{tabular}
\end{center}

Calculate for this test:
\begin{enumerate}
\item The sensibility and the specificity.
\item The positive and negative predictive value.
\item The probability of a correct diagnosis. 
\end{enumerate}

\item We know, from a research study, that 10\% of people over 50 years suffer a particular type or arthritis.
We have developed a new method to detect the disease and after clinical trials we observe that if we apply the method to
people with arthritis we get a positive result in 85\% of cases, while if we apply the method to people without
arthritis, we get a positive result in 4\% of cases.
Answer the following questions:
\begin{enumerate}
\item What is the probability of getting a positive result after applying the method to a random person?
\item If the result of applying the method to one person has been positive, what is the probability of having arthritis?
\end{enumerate}

\item A severe pain without effusion in a particular zone of the knee joint is a symptom of sprained lateral collateral
ligament (SLCL). 
If the sprains in that ligament are classified into grade 1, when there is only distension (60\% of cases), grade 2 when
there is a partial tearing (30\% of cases) and grade 3 when there is a complete tearing (10\% of cases).
Taking into account that the symptom appears in 80\% of cases of grade 1 sprains, 90\% of cases of grade 2 and 100\% of
cases of grade 3, answer the following questions:

\begin{enumerate}
\item If a person has SLCL what is the probability that he or she present severe pain without effusion?
\item What is the diagnosis for a person with severe pain without effusion?
\item From a total of 10000 people with severe pain without effusion, how many are expected to have a grade 1 sprain? 
How many are expected to have a grade 2 sprain? And a grade 3 sprain?
\end{enumerate}


\end{enumerate}

