% Author Alfredo Sánchez Alberca (asalber@ceu.es)

\section{Regression and correlation}
\begin{enumerate}[leftmargin=*,resume]
\item Give some examples of:
\begin{enumerate}
\item Non related variables.
\item Variables that are increasingly related.
\item Variables that are decreasingly related.
\end{enumerate}

\item In an study about the effect of different doses of a medicament, 2 patients got 2 mg and took 5 days to cure, 4
patients got 2 mg and took 6 days to cure, 2 patients got 3 mg ant took 3 days to cure, 4 patients got 3 mg and took 5
days to cure, 1 patient got 3 mg and took 6 days to cure, 5 patients got 4 mg and took 3 days to cure and 2 patients got
4 mg and took 5 days to cure. 
\begin{enumerate}
\item Construct the joint frequency table.
\item Get the marginal frequency distributions and compute the main statistics for every variable. 
\item Compute the covariance and interpret it. 
\end{enumerate}

\item The table below shows the two-dimensional frequency distribution of a sample of 80 persons in a study about the
relation between the blood cholesterol ($X$) in mg/dl and the high blood pressure ($Y$).
\[
\begin{array}{|c||c|c|c||c|}
\hline
X\setminus Y & [110,130) & [130,150) & [150,170) & n_x \\
\hline\hline
[170,190)   &           &     4     &           & 12\\
\hline
[190,210)   &    10     &    12     &     4     &   \\
\hline
[210,230)   &     7     &           &     8     &   \\
\hline
[230,250)   &     1     &           &           & 18\\
\hline\hline
n_y          &           &    30     &    24    &    \\
\hline
\end{array}
\]

\begin{enumerate}
\item Complete the table.
\item Construct the linear regression model of cholesterol on pressure. 
\item Use the linear model to calculate the expected cholesterol for a person with pressure 160 mmHg. 
\item According to the linear model, what is the expected pressure for a person with cholesterol 270 mg/dl?
\end{enumerate}

Use the following sums:
$\sum x=16960$ mg/dl, $\sum y=11160$ mmHg, $\sum x^2=3627200$ (mg/dl)$^2$, $\sum y^2=1576800$ mmHg$^2$ y
$\sum xy=2378800$ mg/dl$\cdot$mmHg.

\item A research study has been conducted to determine the loss of activity of a drug.
The table below shows the results of the experiment.

\begin{center}
\begin{tabular}{|l|r|r|r|r|r|}
\hline
Time (in years) & 1 & 2 & 3 & 4 & 5 \\ \hline
Activity (\%) & 96 & 84 & 70 & 58 & 52 \\ \hline
\end{tabular}
\end{center}

\begin{enumerate}
\item  Construct the linear regression model of activity on time.
\item  According to the linear model, when will the activity be 80\%? When will the drug have lost all activity?
\end{enumerate}

\item A basket team is testing a new stretching program to reduce the injuries during the league. 
The data below show the daily number of minutes doing stretching exercises and the number of injuries along the league. 
\begin{center}
\begin{tabular}{lrrrrrrrr}
\toprule
Stretching minutes & 0 & 30 & 10 & 15 & 5 & 25 & 35 & 40\\
Injuries & 4 & 1 & 2 & 2 & 3 & 1 & 0 & 1\\
\bottomrule
\end{tabular}
\end{center}
\begin{enumerate}
\item Construct the regression line of the number of injuries on the time of stretching. 
\item What is the reduction of injuries for every minute of stretching? 
\item How many minutes of stretching are require for having no injuries? Is reliable this prediction?
\end{enumerate}

Use the following sums ($X$=Number of minutes stretching, and $Y$=number of injuries):
$\sum x_i = 160$ min, $sum y_j=14$ injuries, $\sum x_i^2= 4700$ min$^2$, $\sum y_j^2=36$ injuries$^2$ and $\sum
x_iy_j=160$  min$cdot$injuries.

\item A dietary center is testing a new diet in sample of 12 persons. 
The data below are the number of days of diet and the weight lost (in Kg) until them for every person.  
\begin{center}
(33 , 3.9), (51 , 5.9), (30 , 3.2), (55 , 6.0), (38 , 4.9), (62 , 6.2),\\
(35 , 4.5), (60 , 6.1), (44 , 5.6), (69 , 6.2), (47 , 5.8), (40 , 5.3)
\end{center}
\begin{enumerate}
\item Draw the scatter plot. According to the point cloud, what type of regression model explains better the relation
between the weight lost and the days of diet?
\item Construct the linear regression model and the logarithmic regression model of the weight lost on the number of
days of diet.
\item Use the best model to predict the weight that will lose a person afte 100 days of diet. 
Is this prediction reliable?
\end{enumerate}


\end{enumerate}

